A Logic-Based Framework for Digital Signal Processing

### 2008-06-02 paper/paper.tex
update

\begin{abstract}
In this paper we describe a logic-based framework for Digital Signal Processing. This framework is implemented in SWI-Prolog and develops interfaces to signal processing libraries providing that with modules ``to speak'' DSP. Modules communicate by using common functors, which we use to handle the orchestration of our different analysis components. The system provides the user with a unified and flexible front-end to define signal processing workflows. These workflows involve a number of heterogeneous software libraries. Concurrent Transaction Logic allows us to combine procedural and declarative meaning of queries to execute computations and update data obtaining a powerful query-driven DSP reasoner.

\keywords{DSP predicates, semantic front-end, Concurrent Transaction Logic, computation reasoner, DSP-quering agent}
\end{abstract}

@@

Digital signal processing (DSP) is a broad field and an object of an important research. Withing this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis. Many projects aim at developing formal APIs providing a framework for algorithms implementation and publication. However, these APIs are normally designed as a closed environment which hardly interacts with other related APIs and applications. Although several APIs are wrapped and exported as ``plugin'' libraries which can be embedded in different applications, they still require the implementation of specific and non-interactive hosts.

@@

In this paper we address this problematic situation by proposing and developing SWI-DSP, a logical framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary. We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end. Logic resources are also used to achieve a more capable DSP engine. The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge envolving specific computations.

In \S 2 we identify the most common audio analysis tasks and summarize several relevant \textit{open source} APIs aimed at performing such tasks. We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP. In \S 3 we start viewing the system from the top and how it defines an interesting workflows-based architecture that characterizes input/output relationships between task-oriented modules. These workflows are possible by means of common functorsa and a binary data Standard. \S 4 goes into the implementation of each task-oriented module which are indeed communication layers with dedicated audio processing libraries that are in charge of the actual computation and offer an open development framework.

Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining functors (as common semantic types) and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction. However, this front-end does not present an efficient and manageable framework as it requires specific SWI-Prolog knowledge to create programs suitable for the concrete needs and it does not offer any efficient tabling mechanism. We tackle this problem in \S 5 by defining DSP ``scenarios'' which offer a unified view of the system to the user preserving computational flexibility.

/fixthis
We also describe, as a Standard, how Concurrent Transaction Logic is the solution to provide an efficient engine allowing a formal data updating mechanism and computation reasoning capabilities. Finally, \S 6 presents related works that use SWI-DSP to create an environment for multimedia data management and visualisation. We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.

>>>>

\begin{abstract}
In this paper we describe a logic-based framework for Digital Signal Processing. This framework is implemented as a set of SWI-Prolog modules developing interfaces to heterogeneous signal processing libraries. A set of functors capturing common datatypes and a standard format for binary data are used to mediate our different analysis components. The system provides the user with a unified and flexible front-end to define signal processing workflows. The framework develops a signal processing-enabled reasoner capable of performing the primary tasks through the interfaces: decoding (specific libraries for each format), feature extraction (Vamp plugins), audio processing and effects (LV2/LADSPA plugins) and classification (WEKA machine learning API).

\keywords{Digital Signal Processing, SWI-prolog modules, signal processing workflows, processing-enabled reasoner, interfaced open source engines}
\end{abstract}

@@

Digital signal processing (DSP) is a broad field and an important object of engineering research. Within this research community, particular investigators are discovering and developing new algorithms for signal processing and analysis. Many projects aim at developing formal APIs providing a framework for algorithms implementation. However, these APIs are normally designed as a closed environment which hardly interacts with other related APIs and applications. Although several APIs are wrapped and exported as ``plugin'' libraries which can be embedded in different applications, they still require the implementation of specific and non-interactive hosts.

@@

In this paper we address this problematic situation by proposing and developing SWI-DSP, a logic-based framework for \textit{Audio Processing} (as subset of DSP) that extends the SWI-Prolog vocabulary. We rely on a logic programming language with aggregated audio analysis capabilities getting rid of the application-dependency paradigm and offering a single logic front-end. Logic resources are also used to achieve a more capable DSP engine. The main and primary motivation for a logic-based implementation of such a system is to turn our conception of traditional procedural applications for computation of data towards semantic frameworks to manage knowledge involving specific computations.

In \S 2 we identify the most common audio analysis tasks and summarize several relevant \textit{open source} APIs aimed at performing such tasks. We evaluate the advantages and limitations of these APIs in order to find the requirements for SWI-DSP. In \S 3 we start viewing the system from the top and how it defines an interesting workflows-based architecture that characterizes input/output relationships between task-oriented modules. These workflows are possible by means of common functors and a binary data standard format. \S 4 goes into the implementation of each task-oriented module that develops communication layers with dedicated audio processing libraries that are in charge of the actual computation and offer an open development framework.

Thus, SWI-DSP extends SWI-Prolog vocabulary to perform audio analysis and processing by defining functors (capturing common datatypes) and predicates, triggering external computation performed by dedicated APIs, which are ordered in modules with different levels of abstraction. However, this architecture does not present an efficient and manageable framework as it requires specific module knowledge to create programs suitable for the concrete needs. We tackle this problem in \S 5 by defining an unified and flexible front-end, composed by DSP ``scenarios'', which allows the user to define workflows with understandable semantics.

Data management is an important paradigm involved in many domains and also in the multimedia research domain. Computations can not often be retrieved from session to session or are stored in local filesystems unlikely directed to extend any shareable source of knowledge because they are not expressed in a standard format with common semantics. This problem also encourages the approach presented in this work, based on a predicate calculus implementation (SWI-Prolog), in two different ways:

\begin{itemize}
 \item Concurrent Transcation Logic can be used to extend the predicate calculus to provide an unified logic framework for computations and data updating (given a database implementation).
 \item SWI-Prolog provides an important set of interfaces to other languages (to build user applications embedding the system) and packages that allow us to interact with semantic web technologies which seem to be the best framework to provide a common source of knowledge.
\end{itemize}

Although SWI-DSP does not develop a data management system itself, we discuss this problem in \S 6 and present related works which contribute to achieve such a multimedia knowledge reasoner/manager. We conclude in \S 7 by assessing the value and scope of this work and introducing future advances.
